Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-01 ms.tex
more in 5

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Summary of Numerical Solutions}
\begin{tabular}{lccccccccc}
  \hline
  {$M_{\bullet}$} & {$v_{w}$} & {$r_b^{(b)}$} &  $r_{\rm s}/r_{\rm soi}$ &  {$\left(\frac{\dot{M}}{\dot{M}_{\rm edd}}\frac{0.04}{\eta}\right)_{r_{s}}^{(c)}$} & {$\left(\frac{t_{\rm cool}}{t_{\rm ff}}\frac{0.04}{\eta}\right)_{r_s}^{(d)}$}  \\
 & ($M_{\odot}$) & (km s$^{-1}$) & (pc) & & & & & \\ 
\hline
 & & & $\Gamma^{(a)} = 0.8$ & & & & & \\
  $    10^{ 6 }$ & 300 & 100 & & $ 1.3 \times 10^{ -5 }$ & $ 1.1 \times 10^{ 2 }$ \\
  ... & 600  & 100 & &$ 2.2 \times 10^{ -6 }$ & $ 8.1 \times 10^{ 3 }$ \\
   ...    & 1200  & 100 & &$ 3.8 \times 10^{ -7 }$ & $ 3.7 \times 10^{ 5 }$ \\
  $    10^{ 7 }$ & 300 & 100 & & $ 1.0 \times 10^{ -4 }$ & 6.5 \\
 ...    & ...   & 25 & &$ 7.8 \times 10^{ -5 }$ & 17 \\
 ...      & ...   & 50 & &$ 7.9 \times 10^{ -5 }$ & 12 \\
   ...    & ...   & 200 & &$ 1.2 \times 10^{ -4 }$ & $\mathbf{4.7}$ \\
   ...    & ...   & 400 & &$ 3.6 \times 10^{ -4 }$ & $\mathbf{1.2}$ \\
   ...     & 600  & 100 && $ 7.2 \times 10^{ -6 }$ & $ 2.4 \times 10^{ 3 }$ \\
   ...     & 1200  & 100 && $ 1.3 \times 10^{ -6 }$ & $ 1.2 \times 10^{ 5 }$ \\
  $    10^{ 8 }$ & 300 & 100 & & $ 6.6 \times 10^{ -4 }$ & $\mathbf{1.4}$ \\
...     & ...  & 25 & &$ 1.9 \times 10^{ -4 }$ & 12 \\
   ...     & ...  & 50 & &$ 3.3 \times 10^{ -4 }$ &
  $\mathbf{4.6}$ \\
     ...$^{\dagger}$ & ...  & 50& & $ 9.1 \times 10^{ -4 }$ &
  $\mathbf{5.6}$ \\
   ... & ... & 200 && $ 1.4 \times 10^{ -3 }$ &
  $\mathbf{0.43}$ \\
   ...     & 600  & 100 && $ 2.9 \times 10^{ -5 }$ & $ 5.3 \times 10^{ 2 }$ \\
   ...     & 1200  & 100 && $ 4.0 \times 10^{ -6 }$ & $ 3.8 \times 10^{ 4 }$ \\
  \hline
 & & & $\Gamma = 0.1$ & & & & & & \\
$    10^{ 6 }$ & 600  &25 & & $ 1.6 \times 10^{ -6 }$ & $ 1.6 \times 10^{ 3 }$ \\
  ...  & 600  & 50 & &$ 1.7 \times 10^{ -5 }$ & $\mathbf{33}$ \\
  ...  & 1200  & 100 && $ 5.9 \times 10^{ -8 }$ & $ 1.0 \times 10^{ 6 }$ \\
  $    10^{ 7 }$ & 1200  & 100& & $ 5.1 \times 10^{ -7 }$ & $ 9.4 \times 10^{ 4 }$ \\
  ...  & 600  & 50 && $ 4.5 \times 10^{ -5 }$ & $\mathbf{21}$ \\
  ...  & ...  & 25 & &$ 8.5 \times 10^{ -6 }$ & $ 4.6 \times 10^{ 2 }$ \\
$10^{8}$  & 300  & 50 & & $ 6.0 \times 10^{ -5 }$ & 56 \\
  ...  & 600  & 100 & &$ 8.0 \times 10^{ -4 }$ & $\mathbf{1.6}$ \\
  ...  &  ...  & 25 && $ 2.3 \times 10^{ -5 }$ & $ 2.8 \times 10^{ 2 }$ \\
  ...  &  ...  & 50 && $ 6.0 \times 10^{ -5 }$ & 56 \\
... & 1200 & 100 && $ 2.4 \times 10^{ -6 }$ & $ 2.1 \times 10^{ 4 }$ \\
  ...  & ...  & 50 && $ 2.0 \times 10^{ -6 }$ & $ 3.1 \times 10^{ 4 }$ \\
  ...  &  ...  & 200 && $ 4.2 \times 10^{ -6 }$ & $ 8.4 \times 10^{ 3 }$ \\
  ...  &  ... & 400 && $ 1.2 \times 10^{ -3 }$ & $\mathbf{4.6}$ \\
  \hline
\label{table:models}  
\end{tabular}
\begin{tablenotes}
\item $^{(a)}$ Inner power-law index of stellar Nuker profile.  $^{(b)}$
Break radius of stellar Nuker profile.  $^{(c)}$Accretion rate in
Eddington units, normalized to a stellar mass input parameter $\eta =
0.04$.  $^{(d)}$Ratio of cooling timescale to free-fall timescale at
stagnation radius.  Thermally unstable solutions are shown in bold
text.  $\dagger$Calculated with full radiative cooling included, assuming mass loss parameter $\eta = $? 
\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

Our numerical results, summarized in Table \ref{table:models} and Figure XXX, allow us to study a range of CNM properties and to assess the validity of the analytic estimates from the previous section.  Cases for which $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)< 1$ at the stagnation radius are marked as `thermally unstable', as are cases for which the stagnation radius diverges and hence could not be resolved, resulting in their calculations being terminated.

@@

 Figure~\ref{fig:mdot_mass} shows the SMBH accretion rate for each of our numerical solutions in Eddington units for different values of $\vwO =$ 300, 600 and 1200 km s$^{-1}$, and for both core ($\Gamma$=0.1) and cusp galaxies ($\Gamma$=0.8).  {\bf BDM: say something smart}
%% AG: Describe conservation checks to additionally validate the results.

@@

In Appendix \ref{app:windheat} we calculate the wind heating rate from
stellar winds, $v_{\rm w}^{\star}$, and the mass loss parameter,
$\eta$ (eq.~[\ref{eq:q}]), as a function of age, $\tau_{\star}$, of a
stellar population which is formed impulsively (Fig.~\ref{fig:vwimp}).
At the earliest times ($\tau_{\star} \lesssim 10^{7}$ yr), the wind
heating rate exceeds 1000 km s$^{-1}$, while much later ($\tau_{\star}
\sim t_{\rm h}$) stellar wind heating dominated by main sequence winds
is much lower, $v^{\star}_{\rm w} \sim 50-100 $ km s$^{-1}$.  We show
in $\S\ref{sec:combined}$ that for the case of quasi-continueous star
formation representative of the average star formation history of loss
mass galaxies, the stellar wind heating rate produced by stellar winds
can also be significant, $v_{\rm w}^{\star} \gtrsim 1000$ km s$^{-1}$.
Stellar winds thus contribute a potentially important source of both
energy and mass for the CNM.

@@

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating source (stars, MSPs, SNe Ia, and black hole feedback) calculated for the average star formation history of galaxies containing a given black hole mass.  Again, Ia heating is suppressed in low mass galaxies because the Ia radius exceeds the nominal stagnation radius, with $r_{\rm Ia}/r_{\rm s}$ shown in Fig. ~\ref{fig:eta} (bottom panel).  MSP heating is again found to be negligible for all black hole masses.  The younger stellar populations characterizing galaxies with $\Mbh\lsim 5\times 10^{8} \Msun$ results in stellar wind and Ia heating being dominant, with $v_{w}^{\star} \sim 1000$ km s$^{-1}$ and $v_w^{\rm Ia} \sim 500$ km s$^{-1}$, respectively.  Only for the most massive galaxies $\Mbh\gsim 5\times 10^{8} \Msun$ does the lack of younger stellar populations significantly reduces the role of stellar wind heating.   Solid and dashed lines again show black hole masses for which the flow is thermally stable and thermally unstable, respectively, with $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ as a function of $M_{\bullet}$ shown in the bottom panel.  Strikingly, {\it we find that galactic nuclei experiencing the same average star formation history as their host galaxies produce thermally stable flows across most black hole masses}.  Realistic variations (``burstiness") in the star formation history will result in lower heating rates, as seen in Fig.~X, for which we compare the stellar wind heating rates in the case of sinusoidal star formation pattern...{\bf BDM: say more}.

>>>>

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Summary of Numerical Solutions}
\begin{tabular}{lccccccccc}
  \hline
  {$M_{\bullet}$} & {$v_{w}$} & {$r_b^{(b)}$} &  $r_{\rm s}/r_{\rm soi}$ &  {$\left(\frac{\dot{M}}{\dot{M}_{\rm edd}}\frac{0.04}{\eta}\right)_{r_{s}}^{(c)}$} & {$\left(\frac{t_{\rm cool}}{t_{\rm ff}}\frac{0.04}{\eta}\right)_{r_s}^{(d)}$}  \\
  ($M_{\odot}$) & (km s$^{-1}$) & (pc) &- & - & - &  & \\ 
\hline
 & & & Cusp Galaxies, $\Gamma^{(a)} = 0.8$ & & & & & \\
  $    10^{ 6 }$ & 300 & 100 & & $ 1.3 \times 10^{ -5 }$ & $ 1.1 \times 10^{ 2 }$ \\
  ... & 600  & 100 & &$ 2.2 \times 10^{ -6 }$ & $ 8.1 \times 10^{ 3 }$ \\
   ...    & 1200  & 100 & &$ 3.8 \times 10^{ -7 }$ & $ 3.7 \times 10^{ 5 }$ \\
  $    10^{ 7 }$ & 300 & 100 & & $ 1.0 \times 10^{ -4 }$ & 6.5 \\
 ...    & ...   & 25 & &$ 7.8 \times 10^{ -5 }$ & 17 \\
 ...      & ...   & 50 & &$ 7.9 \times 10^{ -5 }$ & 12 \\
   ...    & ...   & 200 & &$ 1.2 \times 10^{ -4 }$ & $\mathbf{4.7}$ \\
   ...    & ...   & 400 & &$ 3.6 \times 10^{ -4 }$ & $\mathbf{1.2}$ \\
   ...     & 600  & 100 && $ 7.2 \times 10^{ -6 }$ & $ 2.4 \times 10^{ 3 }$ \\
   ...     & 1200  & 100 && $ 1.3 \times 10^{ -6 }$ & $ 1.2 \times 10^{ 5 }$ \\
  $    10^{ 8 }$ & 300 & 100 & & $ 6.6 \times 10^{ -4 }$ & $\mathbf{1.4}$ \\
...     & ...  & 25 & &$ 1.9 \times 10^{ -4 }$ & 12 \\
   ...     & ...  & 50 & &$ 3.3 \times 10^{ -4 }$ &
  $\mathbf{4.6}$ \\
     ...$^{\dagger}$ & ...  & 50& & $ 9.1 \times 10^{ -4 }$ &
  $\mathbf{5.6}$ \\
   ... & ... & 200 && $ 1.4 \times 10^{ -3 }$ &
  $\mathbf{0.43}$ \\
   ...     & 600  & 100 && $ 2.9 \times 10^{ -5 }$ & $ 5.3 \times 10^{ 2 }$ \\
   ...     & 1200  & 100 && $ 4.0 \times 10^{ -6 }$ & $ 3.8 \times 10^{ 4 }$ \\
  \hline
 & & & Core Galaxies, $\Gamma = 0.1$ & & & & & & \\
$    10^{ 6 }$ & 600  &25 & & $ 1.6 \times 10^{ -6 }$ & $ 1.6 \times 10^{ 3 }$ \\
  ...  & 600  & 50 & &$ 1.7 \times 10^{ -5 }$ & $\mathbf{33}$ \\
  ...  & 1200  & 100 && $ 5.9 \times 10^{ -8 }$ & $ 1.0 \times 10^{ 6 }$ \\
  $    10^{ 7 }$ & 1200  & 100& & $ 5.1 \times 10^{ -7 }$ & $ 9.4 \times 10^{ 4 }$ \\
  ...  & 600  & 50 && $ 4.5 \times 10^{ -5 }$ & $\mathbf{21}$ \\
  ...  & ...  & 25 & &$ 8.5 \times 10^{ -6 }$ & $ 4.6 \times 10^{ 2 }$ \\
$10^{8}$  & 300  & 50 & & $ 6.0 \times 10^{ -5 }$ & 56 \\
  ...  & 600  & 100 & &$ 8.0 \times 10^{ -4 }$ & $\mathbf{1.6}$ \\
  ...  &  ...  & 25 && $ 2.3 \times 10^{ -5 }$ & $ 2.8 \times 10^{ 2 }$ \\
  ...  &  ...  & 50 && $ 6.0 \times 10^{ -5 }$ & 56 \\
... & 1200 & 100 && $ 2.4 \times 10^{ -6 }$ & $ 2.1 \times 10^{ 4 }$ \\
  ...  & ...  & 50 && $ 2.0 \times 10^{ -6 }$ & $ 3.1 \times 10^{ 4 }$ \\
  ...  &  ...  & 200 && $ 4.2 \times 10^{ -6 }$ & $ 8.4 \times 10^{ 3 }$ \\
  ...  &  ... & 400 && $ 1.2 \times 10^{ -3 }$ & $\mathbf{4.6}$ \\
  \hline
\label{table:models}  
\end{tabular}
\begin{tablenotes}
\item $^{(a)}$ Inner power-law index of stellar Nuker profile.  $^{(b)}$
Break radius of stellar Nuker profile.  $^{(c)}$Accretion rate in
Eddington units, normalized to a stellar mass input parameter $\eta =
0.04$.  $^{(d)}$Ratio of cooling timescale to free-fall timescale at
stagnation radius.  Thermally unstable solutions are shown in bold
text.  $\dagger$Calculated with full radiative cooling included, assuming mass loss parameter $\eta = $? 
\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

Our numerical results, summarized in Table \ref{table:models} and Figure XXX, allow us to study a range of CNM properties and to assess the validity of the analytic estimates from the previous section.  Cases for which $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|< 1$ at the stagnation radius are marked as `thermally unstable', as are cases for which the stagnation radius diverges and hence could not be resolved, resulting in their calculations being terminated.

@@

 Figure~\ref{fig:mdot_mass} shows the SMBH accretion rate for each of our numerical solutions in Eddington units for different values of $\vwO =$ 300, 600 and 1200 km s$^{-1}$, and for both core ($\Gamma$=0.1) and cusp galaxies ($\Gamma$=0.8).  Shown for comparison is our analytic estimate of $\dot{M}/\dot{M}_{\rm edd}$ from equation (\ref{eq:eddr_analytic}).  For high wind velocities ($v_{w} \gg \sigma$) the stagnation lies well inside the black hole sphere of influence and our analytic estimate provides a good fit to the numerical results.  However, for low wind velocities and/or high $M_{\bullet}$ (large $\sigma$), the numerical accretion rate considerably exceeds the analytic estimate as the stagnation radius diverges to large radii (Fig.~\ref{fig:stag}).  
%% AG: Describe conservation checks to additionally validate the results.

@@

In Appendix \ref{app:windheat} we calculate the wind heating rate from
stellar winds, $v_{\rm w}^{\star}$, and the mass loss parameter,
$\eta$ (eq.~[\ref{eq:q}]), as a function of age, $\tau_{\star}$, of a
stellar population which is formed impulsively (Fig.~\ref{fig:vwImp}).  
At the earliest times ($\tau_{\star} \lesssim 10^{7}$ yr), the wind
heating rate exceeds 1000 km s$^{-1}$, while much later ($\tau_{\star}
\sim t_{\rm h}$) stellar wind heating dominated by main sequence winds
is much lower, $v^{\star}_{\rm w} \sim 50-100 $ km s$^{-1}$.  We show
in $\S\ref{sec:combined}$ that for the case of quasi-continueous star
formation representative of the average star formation history of loss
mass galaxies, the stellar wind heating rate produced by stellar winds
can also be significant, $v_{\rm w}^{\star} \sim 1000$ km s$^{-1}$.  
Stellar winds thus contribute a potentially important source of both
energy and mass for the CNM.  An additional uncertainty in the stellar wind heating rate results from the uncertain thermalization efficiency of massive stellar winds.  We assume this efficiency to be unity, and assumption that is at least partially offset by our conservative assumption of neglecting energy input from core collapse supernovae, which contribute a comparable energy injection rate to massive stellar winds.

@@

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating source (stars, MSPs, SNe Ia, and black hole feedback) calculated for the average star formation history of galaxies containing a given black hole mass.  Again, Ia heating is suppressed in low mass galaxies because the Ia radius exceeds the nominal stagnation radius, with $r_{\rm Ia}/r_{\rm s}$ shown in Fig. ~\ref{fig:eta} (bottom panel).  MSP heating is again found to be negligible for all black hole masses.  The younger stellar populations characterizing galaxies with $\Mbh\lsim 5\times 10^{8} \Msun$ results in stellar wind and Ia heating being dominant, with $v_{w}^{\star} \sim 1000$ km s$^{-1}$ and $v_w^{\rm Ia} \sim 500$ km s$^{-1}$, respectively.  Only for the most massive galaxies $\Mbh\gsim 5\times 10^{8} \Msun$ does the lack of younger stellar populations significantly reduces the role of stellar wind heating.   Solid and dashed lines again show black hole masses for which the flow is thermally stable and thermally unstable, respectively, with $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ as a function of $M_{\bullet}$ shown in the bottom panel.  Strikingly, {\it we find that galactic nuclei experiencing the same average star formation history as their host galaxies produce thermally stable flows across most black hole masses}.

Realistic variations (``burstiness") in the star formation history will result in lower heating rates.  Figure \ref{app:windheat}  shows the heating rate as a function of average black hole mass also for cases in which the current ($z = 0$) star formation is suppressed by a factor $\epsilon_{\rm burst}$ from its average $z = 0$ value for a characteristic timescale $\tau_{\rm var}$ (for details, see Appendix \ref{fig:}).  For low values of $\tau_{\rm var} \lesssim 10^{7}$ yr we see that $v_{w}^{\star}$ is reduced by at most a factor of $\sim 2$ from its average value, even for very small values of $\epsilon_{\rm burst}=$... , while for $\tau_{\rm var} \gtrsim 10^{8}$ yr, wind heating is reduced significantly.  We conclude that variations in the nuclear star formation rate on timescales short compared to the dynamical timescale of the galaxy are unlikely to significantly affect our results.  {\bf BDM: augment/fix this - its just schematic till nick provides his figure}
